---
title: "Assignment 3 - Patrick Torralba"
author: "Patrick Torralba"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

# Preliminary

```{r}
#| label: library
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

library(tidymodels)
library(tidyverse)
library(pROC)
library(kableExtra)

stroke <- read_csv("stroke.csv") |> 
  mutate(stroke = as.factor(stroke))
```

# Question 2

## Need to fix this portion

**Note: this is also part of question 3**

```{r}
#| label: split-data
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

set.seed(999)


stroke_split <-  initial_split(
  stroke, 
  prop = 0.70, 
  strata = stroke)

stroke_train <-  training(stroke_split)

stroke_test <- testing(stroke_split)

```

## Define specs

```{r}
#| label: q2-define-specs
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

log_spec <- logistic_reg() |> 
  set_engine("glm") |> 
  set_mode("classification")

tree_spec <- decision_tree(cost_complexity = 0.001, min_n = 5) |> 
  set_engine("rpart") |> 
  set_mode("classification")


```

## Models

```{r}
#| label: models
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

log_reg_simple <- log_spec |> 
   fit(stroke ~ age + bmi, data = stroke)

log_reg_complex <- log_spec |> 
  fit(stroke ~ age + bmi + smoking_status + heart_disease + hypertension, data = stroke)

tree_simple <- tree_spec |> 
   fit(stroke ~ age + bmi, data = stroke)

tree_complex <- tree_spec |> 
  fit(stroke ~ age + bmi + smoking_status + heart_disease + hypertension, data = stroke)  

```

## Predict Probability & Calculate AUC

```{r}
#| label: AUC
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

get_auc <- function(model, data, model_name) {
  predict(model, new_data = data, type = "prob") |>
    bind_cols(data |> select(stroke)) |>
    roc_auc(
      truth = stroke,
      .pred_1,
      event_level = "second"
    ) |>
    mutate(Model = model_name)
}


```

## Model's AUC

```{r}
#| label: auc-2
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

auc_m1 <- get_auc(log_reg_simple, stroke, "Logistic Regression (age + bmi)")
auc_m2 <- get_auc(log_reg_complex, stroke, "Logistic Regression (Multiple Predictors)")
auc_m3 <- get_auc(tree_simple, stroke, "Decision Tree (age + bmi)")
auc_m4 <- get_auc(tree_complex, stroke, "Decision Tree (Multiple Predictors)")
```

## Comparison table

```{r}
#| label: comp-table
#| message: FALSE
#| warning: FALSE
#| echo: TRUE

auc_comparison <- bind_rows(auc_m1, auc_m2, auc_m3, auc_m4) |>
  select(Model, Metric = .metric, AUC = .estimate)

kable(auc_comparison)

```

# Question 3

## Models

```{r}
#| label: q3-models
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

log_reg_simple_train <- log_spec |> 
  fit(stroke ~ age + bmi, data = stroke_train)

log_reg_complex_train <- log_spec |> 
  fit(
    stroke ~ age + bmi + smoking_status + heart_disease + hypertension,
    data = stroke_train
  )

tree_simple_train <- tree_spec |> 
  fit(stroke ~ age + bmi, data = stroke_train)

tree_complex_train <- tree_spec |> 
  fit(
    stroke ~ age + bmi + smoking_status + heart_disease + hypertension,
    data = stroke_train
  )


predict_log_reg_simple <-  predict(log_reg_simple_train, new_data = stroke_test, type = "prob")


predict_log_reg_complex <- predict(log_reg_complex_train, new_data = stroke_test, type = "prob")


predict_tree_simple <-  predict(tree_simple_train, new_data = stroke_test, type = "prob")


predict_tree_complex <-  predict(tree_complex_train, new_data = stroke_test, type = "prob")


```

## Bind predictions

```{r}
#| label: bind-predictions
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

auc_log_simple <- predict_log_reg_simple |>
  bind_cols(stroke_test |>select(stroke)) |>
  roc_auc(truth = stroke, .pred_1, event_level = "second") |>
  mutate(Model = "Logistic Regression (Simple)", 
         Evaluation = "Test Set")

auc_log_complex <- predict_log_reg_complex |>
  bind_cols(stroke_test |>select(stroke)) |>
  roc_auc(truth = stroke, .pred_1, event_level = "second") |>
  mutate(Model = "Logistic Regression (Complex)", 
         Evaluation = "Test Set")

auc_tree_simple <- predict_tree_simple |>
  bind_cols(stroke_test |>select(stroke)) |>
  roc_auc(truth = stroke, .pred_1, event_level = "second") |>
  mutate(Model = "Decision Tree (Simple)", 
         Evaluation = "Test Set")

auc_tree_complex <- predict_tree_complex |>
  bind_cols(stroke_test |>select(stroke)) |>
  roc_auc(truth = stroke, .pred_1, event_level = "second") |>
  mutate(Model = "Decision Tree (Complex)", 
         Evaluation = "Test Set")
```

## Create AUC Comparison Table

```{r}
#| label: AUC-table-q3
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  


auc_test_table <- bind_rows(auc_log_simple, auc_log_complex, auc_tree_simple, auc_tree_complex) |>  
  select(Model, AUC = .estimate)

kable(auc_test_table)

```

Overfitting in Decision Trees:

The Decision Tree AUCs evaluated on the test set were lower than the AUCs calculated using the full dataset. We expect this to occur because the AUCs from the full dataset essentially "know the answer" already since the model has access to the entire dataset. Whereas, the AUCs on the test set are lower because the model has not seen the test data since it is only trained on the training data.

Interestingly, the logistic regression AUC for both the full dataset and the training dataset changed less dramatically (0.8366 to 00.8317, 0.8477 to 0.841; simple and logistic regression AUC values respectively).