---
title: "Assignment 2 - Patrick Torralba"
author: "Patrick Torralba"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

# Preliminary

```{r}
#| label: import-library
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

library(tidyverse)

stroke <- read_csv("stroke.csv")

```

# Question 1

Define the models

```{r}
#| label: models
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

model1 <- glm(
  data = stroke,
  family = binomial,
  stroke ~ heart_disease + hypertension + smoking_status
)

model2 <- glm(
  data = stroke,
  family = binomial,
  stroke ~ heart_disease + hypertension + smoking_status + age + avg_glucose_level + bmi
)

```

Use the models to obtain probabilities

```{r}
#| label: probability
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

stroke <- stroke |> 
  mutate(
    prob1 = predict(model1, type = "response"),
    prob2 = predict(model2, type = "response"),
    ## I also clean the data for future purposes
    stroke1 = case_when( 
      stroke == 1 ~ "Stroke",
      stroke == 0 ~ "No Stroke",
      .default = NA)
  )

```

Plots

```{r}
#| label: plots
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

## Model 1 Plot
ggplot(stroke, aes(x = prob1, fill = stroke1)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "prob_stroke",
    fill = "obs_stroke"
  )

## Model 2 Plot
ggplot(stroke, aes(x = prob2, fill = stroke1)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "prob_stroke",
    fill = "obs_stroke"
  )
```

Table

```{r}
#| label: table-model
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

library(kableExtra)

stroke_table <- stroke |>
  pivot_longer(
    cols = c(prob1, prob2),
    names_to = "Model",
    values_to = "prob_stroke"
  ) |> 
  mutate(
    Model = case_when(
      Model == "prob1" ~ "Model 1",
      Model == "prob2" ~ "Model 2",
      .default = NA)
  )|> 
  group_by(Model, stroke1) |> 
  summarize(
    "Mean prediction" = mean(prob_stroke),
    "Median prediction" = median(prob_stroke),
    "SD prediction" = sd(prob_stroke),
    .groups = "drop"
  ) |> 
  rename(`Stroke Observed?` = stroke1)


kable(stroke_table)

```

## Answer 1
**It appears that Model 2 is more accurate than Model 1 in terms of predicting stroke outcome. In the summary table, when we look at the median prediction values Model 2 has larger median prediction values than Model 1 for stroke cases only (0.13 vs 0.04, respectively).**

**And, when we look at the density plots, Model 1 has both stroke and no stroke distributions closely overlapping close to zero in contrast to the Model 2 density graph. Therefore, this means that Model 1 is less able to discriminate between stroke outcomes in contrast to Model 2.**

# Question 2

This model includes all covariates except for id and stroke

```{r}
#| label: model-3
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

model3 <- glm(
  data = stroke,
  family = binomial,
  stroke ~ gender + age + hypertension + heart_disease + ever_married + work_type + Residence_type + avg_glucose_level + bmi + smoking_status + stroke
)

stroke_q2 <- stroke |> 
  mutate(
    prob3 = predict(model3, type = "response")
  )


```

Model 3 plot

```{r}
#| label: model-3-plot
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  


## Model 3 Plot
ggplot(stroke_q2, aes(x = prob3, fill = stroke1)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "prob_stroke",
    fill = "obs_stroke"
  )

```

Model 3 Table

```{r}
#| label: table-model-3
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

stroke_table_q2 <- stroke_q2 |>
  pivot_longer(
    cols = c(prob1, prob2, prob3),
    names_to = "Model",
    values_to = "prob_stroke"
  ) |> 
  mutate(
    Model = case_when(
      Model == "prob1" ~ "Model 1",
      Model == "prob2" ~ "Model 2",
      Model == "prob3" ~ "Model 3",
      .default = NA)
  )|> 
  group_by(Model, stroke1) |> 
  summarize(
    "Mean prediction" = mean(prob_stroke),
    "Median prediction" = median(prob_stroke),
    "SD prediction" = sd(prob_stroke),
    .groups = "drop"
  ) |> 
  rename(`Stroke Observed?` = stroke1)


kable(stroke_table_q2)
```

## Answer 2

**No, adding more variables does not appear to improve stroke prediction. Model 3 does have a slightly higher mean prediction (0.14 vs 0.13), but Model 3 also has a lower median stroke prediction value (0.12 vs 0.13, respectively).**

**Likewise, when we compare the density graphs, both Model 3 and Model 2 plots look extremely identical.**

**Therefore, the table and graph mean that adding more variables do not improve stroke prediction outcome. Although the differences between Model 3 and Model 2 appear minimal..**


# Question 3
```{r}
#| label: 50-age-comparison
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

stroke_q3 <- stroke_q2 |> 
  mutate(
    Age = case_when(
      age > 50 ~ "Age > 50",
      .default = "Age <= 50"
    )
  )

```

Plot
```{r}
#| label: q3-plot
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

ggplot(stroke_q3, aes(x = prob3, fill = stroke1)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ Age) +
  labs(
    x = "prob_stroke",
    fill = "obs_stroke"
  )


```


Table
```{r}
#| label: q3-table
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

stroke_table_q3 <- stroke_q3 |>
  group_by(Age, stroke1) |> 
  select(-prob1, -prob2) |> 
  summarize(
    "Mean prediction" = mean(prob3),
    "Median prediction" = median(prob3),
    "SD prediction" = sd(prob3),
    .groups = "drop"
  ) |> 
  rename(`Stroke Observed?` = stroke1)


kable(stroke_table_q3)

```

## Answer 3
**It appears that Model 3 is better at predicting stroke for older individuals because when we look at the Age > 50 plot, there isn't as much overlap in contrast to Age <= 50 denisty plot. This means that Model 3 is better able to discriminate stroke outcomes for older individuals. **

**Also when we look at the summary table, we see that **