---
title: "Assignment 4 - Patrick Torralba"
author: "Patrick Torralba"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

# Preliminary

```{r}
#| label: library
#| message: FALSE
#| warning: FALSE
#| echo: TRUE 

library(tidyverse) #clean data
library(tidymodels)
library(kableExtra) ## create nice data tables

atus <- read_csv("atus.csv")
```

# Before you begin

**I am selecting SCHAPPY as my outcome**

```{r}
#| label: split-data
#| message: FALSE
#| warning: FALSE
#| echo: TRUE 

set.seed(888)

atus_split <- initial_split(
  atus,
  prop = 0.70,
  strata = SCHAPPY
)

atus_train <- training(atus_split)

atus_test <- testing(atus_split)

```

# Question 1

## Define the specs

```{r}
#| label: specs
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lin_reg <- linear_reg() |> 
  set_engine("glm") |> 
  set_mode("regression")
```

## Get all BLS variables
```{r}
#| label: BLS
#| message: FALSE
#| warning: FALSE
#| echo: TRUE

BLS <- atus |> 
  select(starts_with("BLS")) |> 
  select(1:11)

BLS

```


## Define the model

```{r}
#| label: model
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lin_reg_model <- lin_reg |> 
  fit(SCHAPPY ~ BLS_CAREHH +  BLS_CAREHH_ADULT + BLS_CAREHH_KID + BLS_CAREHH_KIDEDUC + BLS_CAREHH_KIDHEALTH + BLS_CAREHH_KIDOTHER + BLS_CAREHH_TRAVEL + BLS_CARENHH + BLS_CARENHH_ADULT + BLS_CARENHH_ADULTHELP,
      data = atus_train)

```


## Predict Probability

```{r}
#| label: predict
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

predict_lin_reg_model <- predict(lin_reg_model, new_data = atus_test) 

## Note to self: only include type = "prob" for classification models

```

## Bind Predictions

```{r}
#| label: bind-predictions
#| message: FALSE
#| warning: FALSE
#| echo: TRUE 

atus_pred <- atus_test |> 
  bind_cols(predict_lin_reg_model)

```

## Compare Actual vs Predicted SCHAPPY Values

```{r}
#| label: compare-viz-happy
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

ggplot(atus_pred, aes(x = SCHAPPY, y = .pred)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(
    x = "Observed SCHAPPY",
    y = "Predicted SCHAPPY",
    title = "Observed vs Predicted Happiness")

```

## SCHAPPY Table

```{r}
#| label: SCHAPPY-table
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

happy_table <- atus_pred |> 
  select(CASEID,BLS_CAREHH, BLS_CAREHH_ADULT,BLS_CAREHH_KID,BLS_CAREHH_KIDEDUC,BLS_CAREHH_KIDHEALTH,BLS_CAREHH_KIDOTHER,BLS_CAREHH_TRAVEL,BLS_CARENHH,BLS_CARENHH_ADULT,BLS_CARENHH_ADULTCARE, BLS_CARENHH_ADULTHELP, SCHAPPY, .pred)

kable(happy_table)
```

## SCHAPPY Coefficients

```{r}
#| label: happy-coefficients
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

coefficients <- lin_reg_model |>
  tidy() |>
  filter(term != "(Intercept)") |>
  mutate(absolute_magnitude = abs(estimate)) |>
  arrange(desc(absolute_magnitude)) |>
  slice_head(n = 10) |>
  select(term, estimate, absolute_magnitude)


kable(coefficients)
```

# Question 2

## Define LASSO spec

```{r}
#| label: lasso-spec
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lasso_spec <- 
  linear_reg(penalty = tune(), mixture = 1) |> 
  set_engine("glmnet")
```

## Locate Best Lambda Value by doing Cross-Validation

```{r}
#| label: locate-lambda-best-value
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

## Define folds
folds <- vfold_cv(atus_train, v = 5)



## Evaluate different values of lambda
lasso_results <- tune_grid(
  lasso_spec,
  SCHAPPY ~ BLS_CAREHH +  BLS_CAREHH_ADULT + BLS_CAREHH_KID + BLS_CAREHH_KIDEDUC + BLS_CAREHH_KIDHEALTH + BLS_CAREHH_KIDOTHER + BLS_CAREHH_TRAVEL + BLS_CARENHH + BLS_CARENHH_ADULT + BLS_CARENHH_ADULTHELP,
  resamples = folds
)

## Find optimal lambda value. This will be the value with the lowest RMSE value.
best_lambda <- select_best(lasso_results, metric = "rmse")

best_lambda_value <- pull(best_lambda, penalty)

```

**The best lambda value is 0.043 because this has the least root mean squared error value. Meaning that when utilized in a model, it is as close as we can get to the observed values.**

## Define LASSO model

```{r}
#| label: lasso-model
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lasso_spec <- 
  linear_reg(penalty = best_lambda_value, mixture = 1) |> 
  set_engine("glmnet")

lasso_model <- lasso_spec |> 
  fit(SCHAPPY ~ BLS_CAREHH +  BLS_CAREHH_ADULT + BLS_CAREHH_KID + BLS_CAREHH_KIDEDUC + BLS_CAREHH_KIDHEALTH + BLS_CAREHH_KIDOTHER + BLS_CAREHH_TRAVEL + BLS_CARENHH + BLS_CARENHH_ADULT + BLS_CARENHH_ADULTHELP,
  data = atus_train)

```

## Predict LASSO Values

```{r}
#| label: predict-lasso
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

predict_lasso_model <- predict(lasso_model, new_data = atus_test) 

## Note to self: only include type = "prob" for classification models

```

## Bind LASSO Predictions

```{r}
#| label: bind-predictions-lasso
#| message: FALSE
#| warning: FALSE
#| echo: TRUE 

lasso_atus_pred <- atus_test |> 
  bind_cols(predict_lasso_model)

```

## Compare Actual vs Predicted LASSO SCHAPPY Values

```{r}
#| label: compare-viz-happy-lasso
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

ggplot(lasso_atus_pred, aes(x = SCHAPPY, y = .pred)) +
  geom_point() +
  labs(
    x = "Observed SCHAPPY",
    y = "Predicted SCHAPPY",
       title = "Observed vs LASSO Predicted Happiness")

```

## LASSO SCHAPPY Coefficients

```{r}
#| label: happy-coefficients-lasso
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lasso_coefficients <- lasso_model |>
  tidy() |>
  filter(term != "(Intercept)", estimate != 0) |>
  mutate(absolute_magnitude = abs(estimate)) |>
  arrange(desc(absolute_magnitude)) |>
  select(term, estimate)


kable(lasso_coefficients)
```

# Question 3

## Compare Coefficients from Linear Regression to Lasso Model

```{r}
#| label: compare-lin-reg-lasso
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  


coefficients_comparison <- full_join(coefficients, lasso_coefficients, by = "term") |> 
  rename(linear_regression_estimate = estimate.x,
         lasso_estimate = estimate.y)

kable(coefficients_comparison)

```

## Calculate and Compare Mean-Squared Errors (MSE)

```{r}
#| label: mse-calculation
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lin_reg_mse <- atus_pred |> 
  summarize(
    MSE = mean((SCHAPPY - .pred)^2)
  )

lin_reg_mse

lasso_mse <- lasso_atus_pred |> 
  summarise(
    MSE = mean((SCHAPPY - .pred)^2)
  )

lasso_mse

```

It looks like the mean-squared errors of both the linear regression and lasso model are very similar based on the displayed output. Therefore, both models have a very similar predictive performance. However, the linear regression model is slightly better based on it's lower RMSE value as depicted in Question 4 below. Also, in terms of mean-squared errors, these differences are unlikely to be scientifically relevant due to their highly identical output. For example, lasso MSE could be 1.44 and linear regression MSE could be 1.435, but due to rounding the output would appear as the same value.

# Question 4

## Define two-stage linear regression model

```{r}
#| label: two-stage-lasso
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lin_reg_model_v2 <- lin_reg |> 
  fit(SCHAPPY ~ BLS_CAREHH_KIDEDUC,
      data = atus_train)

```

## Predictions

```{r}
#| label: two-stage-predictions
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

two_stage_pred <- predict(lin_reg_model_v2, new_data = atus_test)

```

## Bind Predictions to atus_test

```{r}
#| label: two-stage-atus-pred
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

two_stage_atus_pred <- bind_cols(atus_test, two_stage_pred)

```

## Calculate RMSE for two-stage-model

```{r}
#| label: rmse-two-stage
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

two_stage_rmse <- two_stage_atus_pred |> 
  metrics(
    truth = SCHAPPY,
    estimate = .pred
  ) |> 
  filter(.metric == "rmse") |> 
  mutate(model = "two_stage")


two_stage_rmse 

```

## Compare RMSE with all models

```{r}
#| label: rmse-model-compare
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  


lasso_model_rmse <-  lasso_atus_pred |> 
    metrics(
    truth = SCHAPPY,
    estimate = .pred
  ) |> 
  filter(.metric == "rmse") |> 
    mutate(model = "lasso")


lin_reg_model_rmse <- atus_pred |> 
      metrics(
    truth = SCHAPPY,
    estimate = .pred
  ) |> 
  filter(.metric == "rmse") |> 
    mutate(model = "linear_regression")



models_rmse_compare <- bind_rows(lin_reg_model_rmse, lasso_model_rmse, two_stage_rmse)

kable(models_rmse_compare)

```

Since the previous lasso model identified only BLS_CAREHH_KIDEDUC as a non-zero covariate predictor, said predictor was reused when creating the two-stage model. When comparing RMSE values between each model, it appears that all models have the exact same RMSE value, 1.20, after rounding to two decimal places. However, when not rounding, the two-stage model has the lowest RMSE, followed by the linear regression model, and lastly, the lasso model. Therefore, all three models have identical prediction performance with the exception of the two-stage model that performs slightly better due to it's lower RMSE value
