---
title: "Lasso - Patrick Torralba"
author: "Patrick Torralba"
format: 
  html:
    theme: flatly
    fontsize: 0.8em
    code-fold: FALSE
    code-tools: TRUE
    code-copy: TRUE
    toc: TRUE
    toc-depth: 4
    smooth-scroll: TRUE
    embed-resources: TRUE
    fig-align: center
editor: visual
---

# Library

```{r}
#| label: library
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

library(tidyverse)
library(tidymodels)
library(broom)
library(kableExtra)

```

# Import Data

```{r} 
#| label: import
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

flu_train <- read_csv("nhis_flu_train_21_22.csv") |> 
  mutate(
    SHTFLU12M_A = as.factor(case_when(
      str_detect(SHTFLU12M_A, "Yes") ~ 1,
      str_detect(SHTFLU12M_A, "No") ~ 0
    ))) |> 
  slice_sample(n = 10000) |>  
  select(-HHX, -SRVY_YR)

flu_test <- read_csv("nhis_flu_test_23.csv")

flu_predict <- read_csv("nhis_flu_predict_23.csv")

set.seed(123)
```

Note to self: The outcome variable is SHTFLU12M_A and is a binary outcome if the person had a flu vaccine in the past 12 months.

Ideas \* Use LASSO to determine predictors \*

# Step 1

## Define LASSO spec

```{r}
#| label: lasso-specs
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lasso_spec <-
  logistic_reg(penalty = tune(), mixture = 1) |> 
  set_engine("glmnet")

```

## Locate Best Lambda value via Cross-Validation

```{r}
#| label: define-folds
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

## Define folds
folds <- vfold_cv(flu_train, v = 5)
```


## Create preprocessing recipe to remove unknowns, novel, and dummy

```{r}
#| label: preprocessing
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lasso_recipe <- recipe(SHTFLU12M_A ~ ., data = flu_train) |>
  step_unknown(all_nominal_predictors()) |>
  step_novel(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())
```


## Create LASSO workflow

```{r}
#| label: lasso-workflow
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

## Create LASSO workflow
lasso_workflow <- workflow() |>
  add_recipe(lasso_recipe) |>
  add_model(lasso_spec)

```


```{r}
#| label: eval-lambda-values
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  


## Evaluate different values of lambda
lasso_results <- tune_grid(
  lasso_workflow,
  resamples = folds,
  metrics = metric_set(roc_auc)
)


## Find optimal lambda value. This will be the value with the highest ROC AUC value.
best_lambda <- select_best(lasso_results, metric = "roc_auc")

best_lambda_value <- pull(best_lambda, penalty)

```

## Finalize LASSO model

```{r}
#| label: finalize-lasso-model
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

lasso_model <- finalize_workflow(
  lasso_workflow,
  best_lambda
) |> 
  fit(data = flu_train)
```

# Step 2

## Predict LASSO values
```{r}
#| label: predict
#| message: FALSE
#| warning: FALSE
#| echo: TRUE  

flu_pred <- flu_test |> 
  mutate(.pred = predict(lasso_model, new_data = flu_test))

```

